# River basins

```{r load.river, out.width = "100%", dpi = 300, echo = TRUE, warning=FALSE}
library(here)
library(StereoMorph)
library(geomorph)
library(wesanderson)
library(ggplot2)
```

## Read data and define sLMs

```{r read.river, out.width = "100%", dpi = 300, echo = TRUE, warning=FALSE}
shapes <- readShapes("shapes")
shapesGM <- readland.shapes(shapes, nCurvePts = c(10, 10, 10, 10))
```

## Read qualitative data

```{r read qual.river, out.width = "100%", dpi = 300, echo = TRUE, warning=FALSE}
qdata <- read.csv("qdata.csv",
                  header = TRUE,
                  row.names = 1)
```

## Generalised Procrustes Analysis

Landmark data were aligned to a global coordinate system [@RN11622;@RN11623;@RN11563], achieved through generalised Procrustes superimposition [@RN478] performed in R 4.0.4 [@R] using the `geomorph` library v. 3.3.2 [@RN11530;@RN1774]. Procrustes superimposition translates, scales, and rotates the coordinate data to allow for comparisons among objects [@RN11564;@RN478]. The `geomorph` package uses a partial Procrustes superimposition that projects the aligned specimens into tangent space subsequent to alignment in preparation for the use of multivariate methods that assume linear space [@RN1646;@RN11563]. 

```{r gp.river, out.width = "100%", dpi = 300, echo = TRUE, warning=FALSE}
Y.gpa <- gpagen(shapesGM, print.progress = FALSE)
plot(Y.gpa)

# dataframe
gdf <- geomorph.data.frame(shape = Y.gpa$coords,
                            size = Y.gpa$Csize,
                            river = qdata$river)

# add centroid size to qdata
qdata$csz <- Y.gpa$Csize

# revised table of attributes
knitr::kable(qdata,
             align = "lccc",
             caption = "Attributes of Perdiz arrow points included in the sample.")
```

## Boxplot

```{r box.river, out.width = "100%", dpi = 300, echo = TRUE, warning=FALSE}
# attributes
csz <- qdata$csz
river <- qdata$river

# palette
pal = wes_palette("Darjeeling1", 9, type = "continuous")

# boxplot of Perdiz arrow points by river
csz.riv <- ggplot(qdata, aes(x = river, y = csz, color = river)) +
  geom_boxplot() +
  geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.3) +
  scale_color_manual(values = pal) +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(x = 'River Basin', y = 'Centroid Size')
  
# render plot
csz.riv
```

## Principal Components Analysis

Principal components analysis [@RN1746] was used to visualise shape variation among the bottles. The shape changes described by each principal axis are commonly visualised using thin-plate spline warping of a reference 3D mesh [@RN1731;@RN479]. 

```{r pc.river, out.width = "100%", dpi = 300, echo = TRUE, warning=FALSE}
# pca
pca <- gm.prcomp(Y.gpa$coords)
summary(pca)

# set plot parameters
pch.gps <- c(3:6,15:19)[as.factor(river)]
col.gps <- pal[as.factor(river)]
col.hull <- c("#FF0000","#7F5045","#00A08A" ,"#79A645","#F59800","#F2AD00","#F98400","#AAA06B","#5BBCD6")

# pca plot
pc.plot <- plot(pca,
                asp = 1,
                pch = pch.gps,
                col = col.gps)
shapeHulls(pc.plot,
           groups = river,
           group.cols = col.hull)
```

## Procrustes ANOVA - allometry

A residual randomisation permutation procedure (RRPP; n = 10,000 permutations) was used for all Procrustes ANOVAs (Adams and Collyer 2015; Michael L. Collyer and Adams 2018), which has higher statistical power and a greater ability to identify patterns in the data should they be present (Anderson and Ter Braak 2003). To assess whether shape changes differ by size or region (Texas natural geography, as determined by Texas Parks and Wildlife), Procrustes ANOVAs (Goodall 1991) were also run that enlist effect-sizes (zscores) computed as standard deviates of the generated sampling distributions (M. L. Collyer, Sekora, and Adams 2015).

```{r anova.allom, out.width = "100%", dpi = 300, echo = TRUE, warning=FALSE}
# general allometry
fit.size <- procD.lm(shape ~ size,
                     data = gdf,
                     print.progress = FALSE,
                     iter = 9999)

# general allometry
anova(fit.size)
```

## Procrustes ANOVA - shape and size

A residual randomisation permutation procedure (RRPP; n = 10,000 permutations) was used for all Procrustes ANOVAs [@RN1655;@RN11775], which has higher statistical power and a greater ability to identify patterns in the data should they be present [@RN1719]. To assess whether shape changes differ by group (geography and time), Procrustes ANOVAs [@RN1749] were also run that enlist effect-sizes (zscores) computed as standard deviates of the generated sampling distributions [@RN1756].

```{r shape.size.river, out.width = "100%", dpi = 300, echo = TRUE, warning=FALSE}
# size
fit.sz.reg <- procD.lm(size ~ river,
                       data = gdf,
                       print.progress = FALSE,
                       iter = 9999)

# size
anova(fit.sz.reg)

# pairwise comparison of LS means = which differ?
sz.reg <- pairwise(fit.sz.reg,
                   groups = qdata$river)
summary(sz.reg,
        confidence = 0.95,
        test.type = "dist")

# shape
fit.sh.reg <- procD.lm(shape ~ river,
                       data = gdf,
                       print.progress = FALSE,
                       iter = 9999)

# shape
anova(fit.sh.reg)

# pairwise comparison of LS means = which differ?
sh.reg <- pairwise(fit.sh.reg,
                   groups = qdata$river)
summary(sh.reg,
        confidence = 0.95,
        test.type = "dist")
```

## Morphological integration

```{r morph.integ, out.width = "100%", dpi = 300, echo = TRUE, warning=FALSE}
land.gps <- c("A","A","B","A","A","A","A","A","A","A","A","A","B","B","B","B",
              "B","B","B","B","B","B","B","B","B","B","B","B","A","A","A","A",
              "A","A","A","A")
it <- integration.test(Y.gpa$coords,
                       partition.gp = land.gps,
                       iter = 9999,
                       print.progress = FALSE)

summary(it)
plot(it)
```

## Morphological disparity

```{r morph.disp.river, out.width = "100%", dpi = 300, echo = TRUE, warning=FALSE}
morphol.disparity(shape ~ river,
                  groups = qdata$river,
                  data = gdf,
                  print.progress = FALSE,
                  iter = 9999)
```

## Mean shapes

```{r mshape.river, out.width = "100%", dpi = 300, echo = TRUE, warning=FALSE}
# subset landmark coordinates to produce mean shapes
new.coords <- coords.subset(A = Y.gpa$coords,
                            group = qdata$river)
names(new.coords)

# group shape means
mean <- lapply(new.coords, mshape)

# plot mean shapes
plot(mean$Brazos) 
plot(mean$Colorado)
plot(mean$Guadalupe)
plot(mean$Neches)
plot(mean$NRG)
plot(mean$Nueces)
plot(mean$Red)
plot(mean$`San Jacinto`)
plot(mean$Trinity)

# comparison plots
plotRefToTarget(mean$Brazos,
                mean$Colorado,
                method = "points",
                mag = 1)


plotRefToTarget(mean$Brazos, 
                mean$Guadalupe,
                method = "points", 
                mag = 1)


plotRefToTarget(mean$Brazos, 
                mean$Neches, 
                method = "points",
                mag = 1)


plotRefToTarget(mean$Brazos, 
                mean$NRG, 
                method = "points",
                mag = 1)


plotRefToTarget(mean$Brazos, 
                mean$Nueces, 
                method = "points",
                mag = 1)

plotRefToTarget(mean$Brazos, 
                mean$Red, 
                method = "points",
                mag = 1)

plotRefToTarget(mean$Brazos, 
                mean$`San Jacinto`, 
                method = "points",
                mag = 1)

plotRefToTarget(mean$Brazos, 
                mean$Trinity, 
                method = "points",
                mag = 1)

plotRefToTarget(mean$Colorado, 
                mean$Guadalupe, 
                method = "points",
                mag = 1)


plotRefToTarget(mean$Colorado, 
                mean$Neches, 
                method = "points",
                mag = 1)


plotRefToTarget(mean$Colorado,
                mean$NRG, 
                method = "points",
                mag = 1)

plotRefToTarget(mean$Colorado, 
                mean$Nueces, 
                method = "points",
                mag = 1)

plotRefToTarget(mean$Colorado, 
                mean$Red, 
                method = "points",
                mag = 1)

plotRefToTarget(mean$Colorado, 
                mean$`San Jacinto`, 
                method = "points",
                mag = 1)

plotRefToTarget(mean$Colorado, 
                mean$Trinity, 
                method = "points",
                mag = 1)

plotRefToTarget(mean$Guadalupe, 
                mean$Neches, 
                method = "points",
                mag = 1)

plotRefToTarget(mean$Guadalupe, 
                mean$NRG, 
                method = "points",
                mag = 1)

plotRefToTarget(mean$Guadalupe, 
                mean$Nueces, 
                method = "points",
                mag = 1)

plotRefToTarget(mean$Guadalupe, 
                mean$Red, 
                method = "points",
                mag = 1)

plotRefToTarget(mean$Guadalupe, 
                mean$`San Jacinto`, 
                method = "points",
                mag = 1)

plotRefToTarget(mean$Guadalupe, 
                mean$Trinity, 
                method = "points",
                mag = 1)

plotRefToTarget(mean$Neches, 
                mean$NRG, 
                method = "points",
                mag = 1)

plotRefToTarget(mean$Neches, 
                mean$Nueces, 
                method = "points",
                mag = 1)

plotRefToTarget(mean$Neches, 
                mean$Red, 
                method = "points",
                mag = 1)

plotRefToTarget(mean$Neches, 
                mean$`San Jacinto`, 
                method = "points",
                mag = 1)

plotRefToTarget(mean$Neches, 
                mean$Trinity, 
                method = "points",
                mag = 1)

plotRefToTarget(mean$NRG, 
                mean$Nueces, 
                method = "points",
                mag = 1)

plotRefToTarget(mean$NRG, 
                mean$Red, 
                method = "points",
                mag = 1)

plotRefToTarget(mean$NRG, 
                mean$`San Jacinto`, 
                method = "points",
                mag = 1)

plotRefToTarget(mean$NRG, 
                mean$Trinity, 
                method = "points",
                mag = 1)

plotRefToTarget(mean$Red, 
                mean$`San Jacinto`, 
                method = "points",
                mag = 1)

plotRefToTarget(mean$Red, 
                mean$Trinity, 
                method = "points",
                mag = 1)

plotRefToTarget(mean$`San Jacinto`, 
                mean$Trinity, 
                method = "points",
                mag = 1)
```
