--- 
title: "An exploratory analysis of Perdiz arrow point morphology by Texas river basins"
author: "Robert Z. Selden, Jr., Bonnie L. Etter, and B. Sunday Eiselt"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib]
biblio-style: apalike
link-citations: yes
description: "Morphology of Perdiz arrow points in Texas."
---

# Preface {-}

>Reproducibility---the ability to recompute results---and replicability---the chances other experimenters will achieve a consistent result---are two foundational characteristics of successful scientific research [@RN20918].

This volume is written using _Markdown_, and includes all analysis code employed in the study, providing a means for others to reproduce (exactly) those results discussed and expounded upon in the following chapters. The replicable nature of this undertaking provides a means for others to critically assess and evaluate the various analytical components of this study [@RN20915;@RN20916;@RN20917], which is a necessary requirement for the production of reliable knowledge.

## Basis of inquiry

The Toyah Phase has been the center of debate about Texas prehistory, since J. Charles Kelly first defined it in 1947. Known widely as the Toyah Phase Paradox, research has struggled to reconcile the homogenous expression of this protohistoric to historic archaeological record in central Texas and the high levels of ethnic diversity witnessed by French and Spanish explorers at the turn of the 18th-century. Recent studies have related geographic and temporal variation in artifact styles to “communities of practice” using standard 2D measurement techniques, with varying degrees of success. We build on this research by using a 3D geometric morphometric approach to examine previously published photographs of Perdiz projectile points (176 points) from sites across the state, a hallmark of Toyah assemblages. The primary goal of the study is to explore how 2D imaging and geometric morphometric analyses can identify localized variants of Perdiz points based on shape. This work, in turn, may be used to infer patterns of interaction between village clusters through network analyses at the regional scale. Implications for how cultural plurality in central Texas may have influenced Spanish missionization are discussed.

## Primary findings

**Bold text denotes significant differences in _shape_ and _size_.**

* The blade and base of Perdiz arrow points are _morphologically integrated_
* There is _significant allometry_ in the sample
* Perdiz arrow point sizes from the following river basins _differ significantly_:
  + **Brazos and Nueces**
  + Colorado and Neches
  + **Colorado and NRG**
  + **Colorado and Nueces**
  + Colorado and Red
  + **Guadalupe and Nueces**
  + **Neches and Nueces**
  + **NRG and Trinity**
  + Nueces and Trinity
  + Red and Trinity
* Perdiz arrow point shapes from the following river basins _differ significantly_:
  + **Brazos and Nueces**
  + **Colorado and NRG**
  + **Colorado and Nueces**
  + **Guadalupe and Nueces**
  + Neches and NRG
  + **Neches and Nueces**
  + NRG and Nueces
  + NRG and San Jacinto
  + **NRG and Trinity**
  + Nueces and Red

## Funding

Development of analytical components used in this analysis was funded by a grant to RZS (P14AP00138) from the National Center for Preservation Technology and Training.

## Data management

The data and analysis code associated with this project can be accessed through this document or the [GitHub](https://github.com/aksel-blaise/texas.perdiz) repository.

## Colophon

This version of the analysis was generated on `r Sys.time()` using the following computational environment and dependencies: 

```{r colophon, cache = FALSE}
# what R packages and versions were used?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

Current Git commit details are:

```{r}
# where can I find this commit? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```
